
This dissertation would not have been possible without the help and encouragement of many.
I like to thank my adviser, Jonathan May, for his support and patience in shaping me over the past five years.
His right amount of optimism combined with skepticism turned out to be a great gift for me. 
I have benefited from many discussions with my committee members, Shri Narayanan, Xiang Ren, Aiichiro Nakano, Xuezhe Ma, and Chris Mattmann. Huge thanks to Mozhdeh Gheini, Nina Mehrabi, Weiqiu You, Constantine Lignos, and Zhao Zhang, for their fruitful collaborations. 

When I joined USC in 2015 to pursue my MS studies, I had no clue that someday I will be pursuing a Ph.D.
I thank Chris Mattmann for his support and encouragement from the beginning to the end of my graduate school. It is Chris who helped me publish my first paper, took me for an amazing year at NASA JPL where I got motivation to pursue PhD. I thank Kiri Wagstaff for her mentorship and collaboration at JPL; her projects helped me rediscover my passion for NLP. 
I thank Kiri for introducing me to Yolanda Gil, who later introduced me to my future advisor.
Thanks to Peter Zamar, Nanyun Peng, Fred Mostatter and Prem Natarajan for their help with smooth handling of my dual status as a student and a staff at USC. 
Many thanks to Lizsl De Leon for making my dealings with the Computer Science Department an easy affair. I thank USC, as an institution and as a community, for giving me an identity, many opportunities, and resources to become who I am today. I especially thank the USC Employee Tuition Assistance program for funding my doctoral studies. 

My second role as a Research Engineer at USC Information Sciences Institute paved my path to pursuing the work described in this dissertation. I especially like to thank IARPA MATERIAL program and the people behind it. I grew from a newbie to an expert during this program. I thank Scott Miller, Liz Boschee, Joel Barry and Shantanu Agarwal for their productive collaborations. My friends and colleagues at ISI made my Ph.D. journey exciting: Joel Mathew, Joe Mathai, Justin Cho, Katy Felkner,  Meryem M'Hamdi, Nazgol Tavabi, Alex Spangher, Kushal Chawla, Manuel Ciosci, Lukas Ferrer, Micheal Pust, Ashok Deb, Peter Fennel, and Ulf Hermjacob. Many times, I have found answers to my questions either in our group meetings or while having casual conversations over coffee or lunch.

My journey as a first-generation college student from a village in South India to completing a terminal degree at this graduate school in the United States was not a straight road. Firstly, I thank my parents for keeping their promise that I am excused from farm work as long as I perform well in academics! I thank my siblings, Latha and Ambarish, and my wife, Shivani, for their support and encouragement. I thank Mukesh Munisubbanna, Gautham Pai, Monis Khan, Umar Shah, and Rakshith Begane for motivating me to pursue greater goals in my career.